### Analyze Zika infected Glioblastoma data:

---
  output:
  html_document: default
pdf_document: default
---
  ### Finding gene modules that make cells susceptible to Zika infection
  
  ---
  title: "Finding gene modules that make cells susceptible to Zika infection"
output:
  html_document: default
github_document: default
fig_width: 12
fig_height: 4 
---
  
```{r include = FALSE}
### Load data and packages
knitr::opts_chunk$set(fig.width=24, fig.height=8) 
```

Load the required R libraries:
```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}
library(Seurat)
library(Matrix)
library(EnsDb.Hsapiens.v75)
library(rhdf5)
library(dplyr)
```

Prior to this analysis we processed the raw sequencing data with cellranger/STAR aligner and ran an algorithm called CellBender on it to remove ambient RNA.

Here I 1.) load the data 2.) change the gene identifiers from ensemble ids to symbols 3.) construct a metadata matrix (with donor information, Zika exposure etc.)  4.) put the data and metadata into a Seurat object. 5.) load mitochondrial genes from a database for later use 6.) Remove doublets, by loading the doublet scores I got from an algorithm (called scrublet) I ran on the data before.
```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}
setwd('/home/jovyan/HB_ZIK/')
# glioblastoma_counts = read.delim('../data/ZikaGlioblastomas/tic-527/study5953-tic527-star-fc-genecounts.txt', row.names = 1)
# manifest = read.delim('../HB_ZIK/5953stdy_manifest_14517_170919_HB_ZIK_046_048_050_051.csv', sep = ',', skip = 8)
# manifest = manifest[manifest$SUPPLIER.SAMPLE.NAME != "",]
# scrublet_classification = read.delim('scrublet/SmartSeq_predicted_doublets.txt')
# glioblastoma_counts = glioblastoma_counts[,scrublet_classification != 1]
# glioblastoma_metadata = data.frame(matrix('', dim(glioblastoma_counts)[2],4))
# colnames(glioblastoma_metadata) = c('SampleName', 'Technology', 'ZikaExposure', 'Patient')
# glioblastoma_metadata[,1] = unlist(lapply(colnames(glioblastoma_counts), function(x) substring(x,2)))
# glioblastoma_metadata[,2] = rep('SmartSeq', dim(glioblastoma_counts)[2])
# glioblastoma_metadata[,3] = rep('TRUE', dim(glioblastoma_counts)[2])
# glioblastoma_metadata[,4] = substring(manifest$SUPPLIER.SAMPLE.NAME[match(glioblastoma_metadata$SampleName, manifest$SANGER.SAMPLE.ID)], 3, 4)
# patients = c('42','42','43', '43', '45', '45', '46', '46') # (info from sample tracker)
# geneNames = as.matrix(read.table(paste('../data/ZikaGlioblastomas/cellranger302_count_32771_5953STDY855119', as.character(1), '_GRCh38-3_0_0_premrna/filtered_feature_bc_matrix/filtered_feature_bc_matrix/features.tsv', sep = '')))
# 
# geneOccurence = table(geneNames[,2])
# for (gene in names(geneOccurence)){
#   if (geneOccurence[gene] > 1){
#     geneNames[which(geneNames[,2] == gene),2] = paste(geneNames[which(geneNames[,2] == gene),2], '(', geneNames[which(geneNames[,2] == gene),1], ')', sep = '')
#   }
# }
# rownames(glioblastoma_counts) = geneNames[match(rownames(glioblastoma_counts), geneNames[,1]),2]
# glioblastoma_counts = glioblastoma_counts[geneNames[,2],]
# for (i in 1:8){
#   print(i)
#   # prefiltered count_matrix:
#   data_subset <- as.matrix(Read10X_h5(filename = paste('../data/HB_ZIK/HB_ZIK/cellranger302_count_32771_5953STDY855119', as.character(i), '_GRCh38-3_0_0_premrna/output_filtered.h5', sep = ''), use.names = TRUE))
#   scrublet_classification = read.delim(paste('scrublet/5953STDY855119', as.character(i), '_predicted_doublets.txt', sep = ''))
#   print('removed doublets:')
#   print(sum(scrublet_classification))
#   data_subset = data_subset[,scrublet_classification != 1]
#   sampleNames = rownames(data_subset)
#   glioblastoma_counts = cbind(glioblastoma_counts, data_subset)
#   metadata_subset = data.frame(matrix('', dim(data_subset)[2],4))
#   colnames(metadata_subset) = c('SampleName', 'Technology', 'ZikaExposure', 'Patient')
#   metadata_subset[,1] = colnames(data_subset)
#   metadata_subset[,2] = rep('10X', dim(data_subset)[2])
#   metadata_subset[,3] = rep('FALSE', dim(data_subset)[2])
#   metadata_subset[,4] = rep(patients[i], dim(data_subset)[2])
#   glioblastoma_metadata = rbind(glioblastoma_metadata, metadata_subset)
# }
# mitogenes <- genes(EnsDb.Hsapiens.v75, filter = ~ seq_name == "MT")$gene_id
# mitogenes = geneNames[,2][match(mitogenes, geneNames[,1])]
# mitogenes = mitogenes[!is.na(mitogenes)]
# percent.mt = colSums(glioblastoma_counts[rownames(glioblastoma_counts) %in% mitogenes,])/colSums(glioblastoma_counts)
# Glioblastoma <- CreateSeuratObject(glioblastoma_counts, project = 'HB_ZIK', min.cells = 0, min.features = 0)
# Glioblastoma$SampleName = colnames(glioblastoma_counts)
# Glioblastoma$Technology = glioblastoma_metadata$Technology
# Glioblastoma$ZikaExposure = glioblastoma_metadata$ZikaExposure
# Glioblastoma$Patient = glioblastoma_metadata$Patient
# Glioblastoma$percent.mt = percent.mt
# saveRDS(Glioblastoma, file = "../data/ZikaGlioblastomas/zikaGlioblastomas_SeuratObject.rds")
Glioblastoma = readRDS("../data/ZikaGlioblastomas/zikaGlioblastomas_SeuratObject.rds")
```

The QC plots using number of detected genes, number of counts and percent of counts coming from mitochondrial genes (as a proxy for stress), show a couple of outlier cells, which I remove:
  
```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
Glioblastoma.list <- SplitObject(Glioblastoma, split.by = 'Technology')
i = 1
VlnPlot(Glioblastoma.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = 'Patient')
plot1 <- FeatureScatter(Glioblastoma.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Glioblastoma.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Glioblastoma.list[[i]] <- subset(Glioblastoma.list[[i]], subset = nFeature_RNA > 2500 & nFeature_RNA < 12000 & nCount_RNA < 2*10^6 & percent.mt < 0.1)
VlnPlot(Glioblastoma.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = 'Patient')
i = 2
VlnPlot(Glioblastoma.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = 'Patient')
plot1 <- FeatureScatter(Glioblastoma.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Glioblastoma.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Glioblastoma.list[[i]] <- subset(Glioblastoma.list[[i]], subset = nFeature_RNA > 500 & nFeature_RNA < 5000 & nCount_RNA > 0 & nCount_RNA < 4*10^5 & percent.mt < 0.1)
VlnPlot(Glioblastoma.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = 'Patient')
```

The following normalizes, scales and select 2000 particularly variable genes. 

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:2){
Glioblastoma.list[[i]] <- NormalizeData(Glioblastoma.list[[i]], normalization.method = "LogNormalize", scale.factor = 10000)
Glioblastoma.list[[i]] <- FindVariableFeatures(Glioblastoma.list[[i]], selection.method = "vst", nfeatures = 2000)
top25 <- head(VariableFeatures(Glioblastoma.list[[i]]), 25)
plot1 <- VariableFeaturePlot(Glioblastoma.list[[i]])
plot2 <- LabelPoints(plot = plot1, points = top25, repel = TRUE)
CombinePlots(plots = list(plot1, plot2))
all.genes <- rownames(Glioblastoma.list[[i]])
Glioblastoma.list[[i]] <- ScaleData(Glioblastoma.list[[i]], features = all.genes)
}
```

These are the results of the PCA analysis:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:2){
Glioblastoma.list[[i]] <- RunPCA(Glioblastoma.list[[i]], features = VariableFeatures(object = Glioblastoma.list[[i]]))
print(Glioblastoma.list[[i]][["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(Glioblastoma.list[[i]], dims = 1:2, reduction = "pca")
DimPlot(Glioblastoma.list[[i]], reduction = "pca")
DimHeatmap(Glioblastoma.list[[i]], dims = 1, cells = round(dim(Glioblastoma.list[[i]])[2])/2, balanced = TRUE)
DimHeatmap(Glioblastoma.list[[i]], dims = 1:15, cells = round(dim(Glioblastoma.list[[i]])[2])/2, balanced = TRUE)
}
```

Based on the JackStraw procedure I select 14 PCs for further analysis in both cases:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
# for (i in 1:2){
# Glioblastoma.list[[i]] <- JackStraw(Glioblastoma.list[[i]], num.replicate = 100)
# Glioblastoma.list[[i]] <- ScoreJackStraw(Glioblastoma.list[[i]], dims = 1:20)
# JackStrawPlot(Glioblastoma.list[[i]], dims = 1:20)
# ElbowPlot(Glioblastoma.list[[i]], ndims = 40)
# }
```

This is the clustering step:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
n_dimensions = 14
for (i in 1:2){
Glioblastoma.list[[i]] <- FindNeighbors(Glioblastoma.list[[i]], dims = 1:n_dimensions)
Glioblastoma.list[[i]] <- FindClusters(Glioblastoma.list[[i]], resolution = 0.5)
head(Idents(Glioblastoma.list[[i]]), 5)
}
```

Clusters roughly agree with visual seperation on a UMAP plot:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:2){
Glioblastoma.list[[i]] <- RunUMAP(Glioblastoma.list[[i]], dims = 1:n_dimensions)
DimPlot(Glioblastoma.list[[i]], reduction = "umap")
#saveRDS(axonGrowth, file = "axonGrowth_SeuratTutorial.rds")
}
```

Visualize markers and annotate clusters:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
i = 1
#Glioblastoma.0.markers <- FindAllMarkers(Glioblastoma.list[[i]], only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
#saveRDS(Glioblastoma.0.markers, file = "../data/ZikaGlioblastomas/Glioblastoma.0.markers.rds")
Glioblastoma.0.markers = readRDS("../data/ZikaGlioblastomas/Glioblastoma.0.markers.rds")
top = Glioblastoma.0.markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
FeaturePlot(Glioblastoma.list[[i]], features = top$gene)
top10 <- Glioblastoma.0.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
DoHeatmap(Glioblastoma.list[[i]], features = top10$gene, cells = 1:dim(Glioblastoma.list[[i]])[2] + NoLegend())
as.matrix(top10)  
i = 2
#Glioblastoma.1.markers <- FindAllMarkers(Glioblastoma.list[[i]], only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
#saveRDS(Glioblastoma.1.markers, file = "../data/ZikaGlioblastomas/Glioblastoma.1.markers.rds")
Glioblastoma.1.markers = readRDS("../data/ZikaGlioblastomas/Glioblastoma.1.markers.rds")
top = Glioblastoma.1.markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
FeaturePlot(Glioblastoma.list[[i]], features = top$gene)
top5 <- Glioblastoma.1.markers %>% group_by(cluster) %>% top_n(n = 5, wt = avg_logFC)
DoHeatmap(Glioblastoma.list[[i]], features = top5$gene, cells = sample(1:dim(Glioblastoma.list[[i]])[2], 1000))
as.matrix(top10)
```

Vizualize a priori chosen markers of tumour subtypes:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

markers = list(c('DDIT3', 'ENO2', 'VIM', 'ADM', 'LDHA', 'HILPDA'), c('VIM', 'ANXA1', 'ANXA2', 'CHI3L1', 'CD44'), c('CST3', 'GFAP', 'S100B', 'HOPX', 'SLC1A3', 'MLC1'),
               c('PLP1', 'ALCAM', 'OLIG1', 'OMG', 'PLLP'), c('SOX4', 'DCX', 'CD24', 'DLL3', 'SOX11'), c('RND3', 'SOX11', 'DCX', 'CD24', 'STMN4', 'STMN2', 'DLX5', 'DLX6-AS1'))
names(markers) = c('MES-like2', 'MES-like1', 'AC-like', 'OPC-like', 'NPC-like1', 'NPC-like2')
i = 1
Glioblastoma.list[[i]]$MESlike2 = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[1]],])
Glioblastoma.list[[i]]$MESlike1 = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[2]],])
Glioblastoma.list[[i]]$AClike = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[3]],])
Glioblastoma.list[[i]]$OPClike = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[4]],])
Glioblastoma.list[[i]]$NPClike1 = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[5]],])
Glioblastoma.list[[i]]$NPClike2 = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[6]],])
FeaturePlot(Glioblastoma.list[[i]], features = c('MESlike2', 'MESlike1', 'AClike', 'OPClike', 'NPClike1', 'NPClike2'))

i = 2
Glioblastoma.list[[i]]$MESlike2 = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[1]],])
Glioblastoma.list[[i]]$MESlike1 = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[2]],])
Glioblastoma.list[[i]]$AClike = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[3]],])
Glioblastoma.list[[i]]$OPClike = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[4]],])
Glioblastoma.list[[i]]$NPClike1 = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[5]],])
Glioblastoma.list[[i]]$NPClike2 = colSums(Glioblastoma.list[[i]]@assays$RNA@data[markers[[6]],])
FeaturePlot(Glioblastoma.list[[i]], features = c('MESlike2', 'MESlike1', 'AClike', 'OPClike', 'NPClike1', 'NPClike2'))
```

Vizualize donor:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
i = 1
DimPlot(Glioblastoma.list[[i]], reduction = "umap", group.by = 'Patient')
i = 2
DimPlot(Glioblastoma.list[[i]], reduction = "umap", group.by = 'Patient')
```

Visualize markers and annotate clusters:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
i = 1
#Glioblastoma.0.markers <- FindAllMarkers(Glioblastoma.list[[i]], only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
#saveRDS(Glioblastoma.0.markers, file = "../data/ZikaGlioblastomas/Glioblastoma.0.markers.rds")
Glioblastoma.0.markers = readRDS("../data/ZikaGlioblastomas/Glioblastoma.0.markers.rds")
top = Glioblastoma.0.markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
FeaturePlot(Glioblastoma.list[[i]], features = top$gene)
top10 <- Glioblastoma.0.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
DoHeatmap(Glioblastoma.list[[i]], features = top10$gene, cells = 1:dim(Glioblastoma.list[[i]])[2] + NoLegend())
as.matrix(top10)  
i = 2
#Glioblastoma.1.markers <- FindAllMarkers(Glioblastoma.list[[i]], only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
#saveRDS(Glioblastoma.1.markers, file = "../data/ZikaGlioblastomas/Glioblastoma.1.markers.rds")
Glioblastoma.1.markers = readRDS("../data/ZikaGlioblastomas/Glioblastoma.1.markers.rds")
top = Glioblastoma.1.markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
FeaturePlot(Glioblastoma.list[[i]], features = top$gene)
top5 <- Glioblastoma.1.markers %>% group_by(cluster) %>% top_n(n = 5, wt = avg_logFC)
DoHeatmap(Glioblastoma.list[[i]], features = top5$gene, cells = sample(1:dim(Glioblastoma.list[[i]])[2], 1000))
as.matrix(top10)
```

Vizualize a priori chosen markers of tumour subtypes:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
viral_response = c('DDIT3', 'IFNB1')
i = 1
Glioblastoma.list[[i]]$viral_response = colSums(Glioblastoma.list[[i]]@assays$RNA@data[viral_response,])
FeaturePlot(Glioblastoma.list[[i]], features = 'viral_response')
i = 2
Glioblastoma.list[[i]]$viral_response = colSums(Glioblastoma.list[[i]]@assays$RNA@data[viral_response,])
FeaturePlot(Glioblastoma.list[[i]], features = 'viral_response')
```


