### Analyze Zika infected Glioblastoma data:

---
  output:
  html_document: default
pdf_document: default
---
  ### Finding gene modules that make cells susceptible to Zika infection
  
  ---
  title: "Finding gene modules that make cells susceptible to Zika infection"
output:
  html_document: default
github_document: default
fig_width: 12
fig_height: 4 
---
  
```{r include = FALSE}
### Load data and packages
knitr::opts_chunk$set(fig.width=24, fig.height=8) 
```

Load the required R libraries:
```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}
library(Seurat)
library(Matrix)
library(EnsDb.Hsapiens.v75)
library(rhdf5)
library(dplyr)
library(ggplot2)
```

Prior to this analysis we processed the raw sequencing data with cellranger/STAR aligner and ran an algorithm called CellBender on it to remove ambient RNA.

Here I 1.) load the data 2.) change the gene identifiers from ensemble ids to symbols 3.) construct a metadata matrix (with donor information, Zika exposure etc.)  4.) put the data and metadata into a Seurat object. 5.) load mitochondrial genes from a database for later use 6.) Remove doublets, by loading the doublet scores I got from an algorithm (called scrublet) I ran on the data before.

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}
setwd('/home/jovyan/HB_ZIK/')
# glioblastoma_counts = read.delim('../data/HB_ZIK/HB_ZIK/study5953-zika-star-genecounts.txt', row.names = 1)
# glioblastoma_counts = glioblastoma_counts[!rownames(glioblastoma_counts) %in% c("N_ambiguous", "N_multimapping" , "N_noFeature" ,
#                                                                                 "N_unmapped"),]
# # glioblastoma_counts = read.delim('../data/ZikaGlioblastomas/tic-527/study5953-tic527-star-fc-genecounts.txt', row.names = 1)
# manifest1 = read.delim('../HB_ZIK/5953stdy_manifest_14517_170919_HB_ZIK_046_048_050_051.csv', sep = ',', skip = 8)
# manifest2 = read.delim('../HB_ZIK/5953stdy_manifest_14797_111119_HB_ZIK_01.02.2020_updated.csv', sep = ',', skip = 8)
# manifest1 = manifest1[manifest1$SUPPLIER.SAMPLE.NAME != "",]
# manifest2 = manifest2[manifest2$SUPPLIER.SAMPLE.NAME != "",]
# keepCols = colnames(manifest2)[colnames(manifest2) %in% colnames(manifest1)]
# manifest1 = manifest1[,keepCols]
# manifest2 = manifest2[,keepCols]
# manifest = rbind(manifest1, manifest2)
# scrublet_classification = read.delim('scrublet/SmartSeq_predicted_doublets.txt')
# glioblastoma_counts = glioblastoma_counts[,scrublet_classification != 1]
# glioblastoma_metadata = data.frame(matrix('', dim(glioblastoma_counts)[2],4))
# colnames(glioblastoma_metadata) = c('SampleName', 'Technology', 'ZikaExposure', 'Patient')
# glioblastoma_metadata[,1] = unlist(lapply(colnames(glioblastoma_counts), function(x) substring(x,2)))
# glioblastoma_metadata[,2] = rep('SmartSeq', dim(glioblastoma_counts)[2])
# glioblastoma_metadata[,3] = rep('TRUE', dim(glioblastoma_counts)[2])
# glioblastoma_metadata[,4] = substring(manifest$SUPPLIER.SAMPLE.NAME[match(glioblastoma_metadata$SampleName, manifest$SANGER.SAMPLE.ID)], 3, 4)
# patients = c('42','42','43', '43', '45', '45', '46', '46') # (info from sample tracker)
# geneNames = as.matrix(read.table(paste('/home/jovyan/data/HB_ZIK/HB_ZIK/cellranger302_count_32771_5953STDY855119', as.character(1), '_GRCh38-3_0_0_premrna/filtered_feature_bc_matrix/features.tsv', sep = '')))
# geneNames = rbind(geneNames, c("zikamcherryG", "zikamcherryG", 'Gene', 'Expression'))
# 
# geneOccurence = table(geneNames[,2])
# for (gene in names(geneOccurence)){
#   if (geneOccurence[gene] > 1){
#     geneNames[which(geneNames[,2] == gene),2] = paste(geneNames[which(geneNames[,2] == gene),2], '(', geneNames[which(geneNames[,2] == gene),1], ')', sep = '')
#   }
# }
# 
# rownames(glioblastoma_counts) = geneNames[match(rownames(glioblastoma_counts), geneNames[,1]),2]
# 
# glioblastoma_counts = glioblastoma_counts[geneNames[,2],]
# for (i in 1:8){
#   print(i)
#   # prefiltered count_matrix:
#   data_subset <- as.matrix(Read10X_h5(filename = paste('../data/HB_ZIK/HB_ZIK/cellranger302_count_32771_5953STDY855119', as.character(i), '_GRCh38-3_0_0_premrna/output_filtered.h5', sep = ''), use.names = TRUE))
#   scrublet_classification = read.delim(paste('scrublet/5953STDY855119', as.character(i), '_predicted_doublets.txt', sep = ''))
#   print('removed doublets:')
#   print(sum(scrublet_classification))
#   data_subset = data_subset[,scrublet_classification != 1]
#   sampleNames = rownames(data_subset)
#   data_subset = rbind(data_subset, rep(0,dim(data_subset)[2]))
#   rownames(data_subset)[length(rownames(data_subset))] = "zikamcherryG"
#   glioblastoma_counts = cbind(glioblastoma_counts, data_subset)
#   metadata_subset = data.frame(matrix('', dim(data_subset)[2],4))
#   colnames(metadata_subset) = c('SampleName', 'Technology', 'ZikaExposure', 'Patient')
#   metadata_subset[,1] = colnames(data_subset)
#   metadata_subset[,2] = rep('10X', dim(data_subset)[2])
#   metadata_subset[,3] = rep('FALSE', dim(data_subset)[2])
#   metadata_subset[,4] = rep(patients[i], dim(data_subset)[2])
#   glioblastoma_metadata = rbind(glioblastoma_metadata, metadata_subset)
# }
# mitogenes <- genes(EnsDb.Hsapiens.v75, filter = ~ seq_name == "MT")$gene_id
# mitogenes = geneNames[,2][match(mitogenes, geneNames[,1])]
# mitogenes = mitogenes[!is.na(mitogenes)]
# percent.mt = colSums(glioblastoma_counts[rownames(glioblastoma_counts) %in% mitogenes,])/colSums(glioblastoma_counts)
# Glioblastoma <- CreateSeuratObject(glioblastoma_counts, project = 'HB_ZIK', min.cells = 0, min.features = 0)
# Glioblastoma$SampleName = colnames(glioblastoma_counts)
# Glioblastoma$Technology = glioblastoma_metadata$Technology
# Glioblastoma$ZikaExposure = glioblastoma_metadata$ZikaExposure
# Glioblastoma$Patient = glioblastoma_metadata$Patient
# Glioblastoma$percent.mt = percent.mt
# saveRDS(Glioblastoma, file = "../data/HB_ZIK/HB_ZIK/zikaGlioblastomas_SeuratObject.rds")
Glioblastoma = readRDS("../data/jhubData/HB_ZIK/HB_ZIK/zikaGlioblastomas_SeuratObject.rds")
# Glioblastoma = Glioblastoma[,!is.na(Glioblastoma$Patient)]
# Glioblastoma = Glioblastoma[,Glioblastoma$Patient != '46']
```

The QC plots using number of detected genes, number of counts and percent of counts coming from mitochondrial genes (as a proxy for stress), show a couple of outlier cells, which I remove:
  
```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
Glioblastoma.list <- SplitObject(Glioblastoma, split.by = 'Technology')
i = 1
VlnPlot(Glioblastoma.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = 'Patient')
plot1 <- FeatureScatter(Glioblastoma.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Glioblastoma.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Glioblastoma.list[[i]] <- subset(Glioblastoma.list[[i]], subset = nFeature_RNA > 2500 & nFeature_RNA < 12000 & nCount_RNA < 2*10^6 & percent.mt < 0.1)
VlnPlot(Glioblastoma.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = 'Patient')
i = 2
VlnPlot(Glioblastoma.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = 'Patient')
plot1 <- FeatureScatter(Glioblastoma.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Glioblastoma.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Glioblastoma.list[[i]] <- subset(Glioblastoma.list[[i]], subset = nFeature_RNA > 500 & nFeature_RNA < 5000 & nCount_RNA > 0 & nCount_RNA < 4*10^5 & percent.mt < 0.1)
VlnPlot(Glioblastoma.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = 'Patient')
```

Integrate 10X data from different donors:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
n_dimensions = 15
allGenes = rownames(Glioblastoma.list[[2]])
Glioblastoma.10x = SplitObject(Glioblastoma.list[[2]], split.by = 'Patient')
glioblastoma.anchors <- FindIntegrationAnchors(object.list = Glioblastoma.10x, dims = 1:n_dimensions)
glioblastoma.10x.integrated <- IntegrateData(anchorset = glioblastoma.anchors, dims = 1:n_dimensions, features.to.integrate = allGenes)
DefaultAssay(glioblastoma.10x.integrated) <- "integrated"
glioblastoma.10x.integrated = FindVariableFeatures(glioblastoma.10x.integrated, selection.method = "vst", nfeatures = 2000)
glioblastoma.10x.integrated <- ScaleData(glioblastoma.10x.integrated, verbose = FALSE, features = allGenes)
glioblastoma.10x.integrated <- RunPCA(glioblastoma.10x.integrated, npcs = n_dimensions, verbose = FALSE)
glioblastoma.10x.integrated <- RunUMAP(glioblastoma.10x.integrated, reduction = "pca", dims = 1:n_dimensions)
DimPlot(glioblastoma.10x.integrated, reduction = "umap", group.by = "Patient")
```

Calculate a module score for each cell using the method in the Neftel 2019 paper:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

# Load Neftel modules:

options(stringsAsFactors = FALSE)
modules = read.delim('/home/jovyan/HB_ZIK/1-s2.0-S0092867419306877-mmc2.csv', skip = 4, sep = ',')

MESlike = c(modules['MES2'][modules['MES2'] != ''], modules['MES1'][modules['MES1'] != ''])
AClike = modules['AC'][modules['AC'] != '']
OPClike = modules['OPC'][modules['OPC'] != '']
NPClike = c(modules['NPC2'][modules['NPC2'] != ''], modules['NPC1'][modules['NPC1'] != ''])
 
# Again to be consistent with the Neftel paper note which genes have more than 4 average log2 counts:

normCounts = 2^(as.matrix(GetAssayData(object = glioblastoma.10x.integrated, slot = "data")))
             
average = log2(rowMeans(normCounts))
keepGenes = average > 4
rm(average, normCounts)

# Define relative expression:

meanExpression = rowMeans(as.matrix(GetAssayData(object = glioblastoma.10x.integrated, slot = "data")))[keepGenes]

relativeExpression = as.matrix(GetAssayData(object = glioblastoma.10x.integrated, slot = "data"))[keepGenes,]-meanExpression 

# Bin genes into 30 groups based on expression:
b = seq(from = min(meanExpression)-0.0001, to = max(meanExpression)+0.0001, length.out = 30)
bins = .bincode(meanExpression, b, TRUE)
names(bins) = names(meanExpression)

# Subset signatures to include only available genes:
MESlike = MESlike[MESlike %in% names(meanExpression)]
AClike = AClike[AClike %in% names(meanExpression)]
OPClike = OPClike[OPClike %in% names(meanExpression)]
NPClike = NPClike[NPClike %in% names(meanExpression)]

# Extract control gene set for each signature:
MESlike_bins = bins[MESlike]
AClike_bins = bins[AClike]
OPClike_bins = bins[OPClike]
NPClike_bins = bins[NPClike]

MESlike_control = unname(unlist(lapply(MESlike_bins, function(b) sample(names(bins[bins == b]), 100, replace = T))))
AClike_control = unname(unlist(lapply(AClike_bins, function(b) sample(names(bins[bins == b]), 100, replace = T))))
OPClike_control = unname(unlist(lapply(OPClike_bins, function(b) sample(names(bins[bins == b]), 100, replace = T))))
NPClike_control = unname(unlist(lapply(NPClike_bins, function(b) sample(names(bins[bins == b]), 100, replace = T))))

# Calculate score:

MESlike_score = colMeans(relativeExpression[MESlike,]) - colMeans(relativeExpression[MESlike_control,]) 
AClike_score = colMeans(relativeExpression[AClike,]) - colMeans(relativeExpression[AClike_control,]) 
OPClike_score = colMeans(relativeExpression[OPClike,]) - colMeans(relativeExpression[OPClike_control,]) 
NPClike_score = colMeans(relativeExpression[NPClike,]) - colMeans(relativeExpression[NPClike_control,]) 

# Add to object:

glioblastoma.10x.integrated$MESlike_score = MESlike_score
glioblastoma.10x.integrated$AClike_score = AClike_score
glioblastoma.10x.integrated$OPClike_score = OPClike_score
glioblastoma.10x.integrated$NPClike_score = NPClike_score

```

Visualize distribution of scores in 10x data:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

glioblastoma.10x.integrated.patients = SplitObject(glioblastoma.10x.integrated, split.by = 'Patient')

# Make paired plots for donors 45 and 46:

par(mfrow=c(3,4))

# 45

hist(glioblastoma.10x.integrated.patients$`45`$MESlike_score, col='skyblue', border=F, freq = F, xlab = 'MESlike score', main = 'Patient 45')

hist(glioblastoma.10x.integrated.patients$`45`$OPClike_score, col='skyblue',border=F, freq = F, xlab = 'OPClike score', main = 'Patient 45')

hist(glioblastoma.10x.integrated.patients$`45`$AClike_score, col='skyblue',border=F, freq = F, xlab = 'AClike score', main = 'Patient 45')

hist(glioblastoma.10x.integrated.patients$`45`$NPClike_score, col='skyblue',border=F, freq = F, xlab = 'NPClike score', main = 'Patient 45')

# 46

hist(glioblastoma.10x.integrated.patients$`46`$MESlike_score, col='skyblue',border=F, freq = F, xlab = 'MESlike score', main = 'Patient 46')

hist(glioblastoma.10x.integrated.patients$`46`$OPClike_score, col='skyblue',border=F, freq = F, xlab = 'OPClike score', main = 'Patient 46')

hist(glioblastoma.10x.integrated.patients$`46`$AClike_score, col='skyblue',border=F, freq = F, xlab = 'AClike score', main = 'Patient 46')

hist(glioblastoma.10x.integrated.patients$`46`$NPClike_score, col='skyblue',border=F, freq = F, xlab = 'NPClike score', main = 'Patient 46')

# 42,43,45,46 in 10x vs 48,50,51,54 in SmartSeq

hist(c(glioblastoma.10x.integrated.patients$`42`$MESlike_score, glioblastoma.10x.integrated.patients$`43`$MESlike_score, glioblastoma.10x.integrated.patients$`45`$MESlike_score, glioblastoma.10x.integrated.patients$`46`$MESlike_score), col='skyblue',border=F, freq = F, xlab = 'MESlike score', main = 'Patients 42,43,45,46 vs 48,50,51,54')

hist(c(glioblastoma.10x.integrated.patients$`42`$OPClike_score, glioblastoma.10x.integrated.patients$`43`$OPClike_score,glioblastoma.10x.integrated.patients$`45`$OPClike_score, glioblastoma.10x.integrated.patients$`46`$OPClike_score), col='skyblue',border=F, freq = F, xlab = 'OPClike score', main = 'Patients 42,43,45,46 vs 48,50,51,54')

hist(c(glioblastoma.10x.integrated.patients$`42`$AClike_score, glioblastoma.10x.integrated.patients$`43`$AClike_score, glioblastoma.10x.integrated.patients$`45`$AClike_score, glioblastoma.10x.integrated.patients$`46`$AClike_score), col='skyblue',border=F, freq = F, xlab = 'AClike score', main = 'Patients 42,43,45,46 vs 48,50,51,54')

hist(c(glioblastoma.10x.integrated.patients$`42`$NPClike_score, glioblastoma.10x.integrated.patients$`43`$NPClike_score, glioblastoma.10x.integrated.patients$`45`$NPClike_score, glioblastoma.10x.integrated.patients$`46`$NPClike_score), col='skyblue',border=F, freq = F, xlab = 'NPClike score', main = 'Patients 42,43,45,46 vs 48,50,51,54')

```

Cluster data:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

glioblastoma.10x.integrated <- FindNeighbors(glioblastoma.10x.integrated, dims = 1:n_dimensions)
glioblastoma.10x.integrated <- FindClusters(glioblastoma.10x.integrated, resolution = 0.5)

```

Define markers of brain cell types and add scores to scale data slot:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

cell.markers = c(
  'Neurons' =  'RBFOX3', # NeuN
  'Immature Neurons' = 'DCX',
  'NSC.1' = 'EMX2', # Apparently also in astrocytes
  'NSC.2' = 'PAX6',
  'Activated neurons' = 'FOS', 

  'NSC and Astrocytes' = 'GFAP',

  'VGLUT1 excitatory' = 'SLC17A7',
  'VGLUT2 excitatory' = 'SLC17A6',

  'Layer I inhibitory' = 'RELN',
  'Layer II excitatory' = 'LAMP5',
  'Layer II/III-IV excitatory.1' = 'CUX1',
  'Layer II/III-IV excitatory.2' = 'CUX2',
  'Layer IV excitatory' = 'RORB',
  'Layer Va excitatory' = 'PCP4',
  'Layer Vb excitatory' = 'HTR2C',
  'Layer VIa excitatory' = 'OPRK1',
  'Layer VIb excitatory' = 'NR4A2',

  'HPA Upper layer' = 'TPPP3',
  'HPA Mid layer' = 'NECAB1',
  'HPA Lower layer' = 'PCP4',

  'GABAergic' = 'GAD1',
  'PVALB interneuron' = 'PVALB',
  'SST interneuron' = 'SST',
  '5HT3aR interneuron' = 'HTR3A',
  'VIP interneuron' = 'VIP', #Subset of 5HTraR

  'Cholinergic' = 'ACHE',
  'Dopaminergic' = 'TH',
  'Serotinergic' = 'TPH1',

  'Astrocytes' = 'AQP4',

  'Oligodendrocytes.1' = 'PLP1',
  'Oligodendrocytes.1' = 'MOG',

  'OPC' = 'PDGFRA',

  'Endothelial cells' = 'APOLD1',

  'Microglia.1' = 'CCL4',
  'Microglia.2' = 'CCL3',
  'Microglia.3' = 'TMEM119',

  'Microglia and astrocytes' = 'CX3CR1',

  'pre-Bötzinger interneurons' = 'TACR1',
  'T cells' = 'CD3G'
  )

kriegstein.markers = c(
'Mature IPC/Newborn Neuron'		=	'ADCYAP1',
'Glycolytic Progenitor'		=	'BHLHE22',
'OPC'		=	'CHRNA3',
'Oligodendrocyte'		=	'SCRT1',
'Microglia'		=	'CALB2',
'Mixed Progenitor/Neuron'		=	'TAC3',
'Endothelial'		=	'SCRT2',
'Tumor Associated Macrophage'		=	'CRABP1',
'Oligodendrocyte'		=	'PPP1R17',
'Pericyte'		= 'DOC2B',
'OPC'		=	'CABP7',
'Oligodendrocyte'		=	'NHLH2',
'Endothelial'		=	'CDH7',
'Dividing Neuron'		=	'DLX6',
'Protoplasmic Astrocyte'		=	'MYT1L',
'B Cells'		=	'NHLH1',
'Microglia'		=	'TP73',
'Dividing Neuron'		=	'NEUROD4',
'Radial Glia'		=	'OPRD1',
'Endothelial'		=	'VIPR2',
'Pericyte'		=	'PTPN5',
'Endothelial'		=	'PCSK2',
'Radial Glia'		=	'CAMKV',
'Oligodendrocyte'		=	'ACTL6B',
'Dividing OPC'		=	'DACH1',
'Radial Glia'		=	'FNDC5',
'Radial Glia'		=	'KLHL35',
'Radial Glia'		=	'ELAVL2',
'Microglia'		=	'ONECUT2',
'Dividing Neuron'		=	'SYT4',
'OPC'		=	'SRRM4',
'Microglia'		=	'SH3GL2',
'Microglia'		=	'SVEP1',
'Glycolytic Progenitor'		=	'CPLX1',
'Unknown'		=	'RBFOX1',
'Microglia'		=	'L1CAM',
'Microglia'		=	'CELF3',
'Radial Glia'		=	'DLX5',
'OPC'		=	'ST8SIA2',
'Dividing Neuron'		=	'TSPAN18',
'Immature Astrocyte'		=	'LHX1',
'Mixed Progenitor/Neuron'		=	'BRSK2',
'Immature Astrocyte'		=	'DUSP26',
'Microglia'		=	'SRRM3',
'Neuron'		=	'NKAIN1',
'Red blood cells'		=	'GAL',
'CGE iN'		=	'GRIN2B',
'Microglia'		=	'RIMS3',
'Dividing B Cells'		=	'CELSR3',
'Oligodendrocyte'		=	'SYT7',
'Dividing Progenitor'		=	'HUNK',
'Dividing Progenitor'		=	'ST8SIA3'
)

glioblastoma.10x.integrated@assays$integrated@scale.data = rbind(glioblastoma.10x.integrated$MESlike_score, glioblastoma.10x.integrated@assays$integrated@scale.data)
rownames(glioblastoma.10x.integrated@assays$integrated@scale.data)[1] = 'MESlike'

glioblastoma.10x.integrated@assays$integrated@scale.data = rbind(glioblastoma.10x.integrated$OPClike_score, glioblastoma.10x.integrated@assays$integrated@scale.data)
rownames(glioblastoma.10x.integrated@assays$integrated@scale.data)[1] = 'OPClike'

glioblastoma.10x.integrated@assays$integrated@scale.data = rbind(glioblastoma.10x.integrated$NPClike_score, glioblastoma.10x.integrated@assays$integrated@scale.data)
rownames(glioblastoma.10x.integrated@assays$integrated@scale.data)[1] = 'NPClike'

glioblastoma.10x.integrated@assays$integrated@scale.data = rbind(glioblastoma.10x.integrated$AClike_score, glioblastoma.10x.integrated@assays$integrated@scale.data)
rownames(glioblastoma.10x.integrated@assays$integrated@scale.data)[1] = 'AClike'

```

Visualize clusters and markers in a heatmap:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
pdf('ClusterHeatmap_NeftelScores.pdf', width = 20, height = 10)
print(DoHeatmap(glioblastoma.10x.integrated, features = c('MESlike', 'AClike', 'OPClike', 'NPClike'), slot = 'scale.data',
                group.by = 'Patient') + NoLegend())
dev.off()

pdf('ClusterHeatmap_StandardMarkers.pdf', width = 20, height = 10)
print(DoHeatmap(glioblastoma.10x.integrated, features = unlist(cell.markers), slot = 'scale.data') + NoLegend()) 
dev.off()

pdf('ClusterHeatmap_KriegsteinMarkers.pdf', width = 20, height = 10)
print(DoHeatmap(glioblastoma.10x.integrated, features = unlist(kriegstein.markers), slot = 'scale.data') + NoLegend())
dev.off()
```

Assign identity to cluster based on marker expression:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
new.cluster.ids <- list(
  
                    c("OPClike",
                     "NPClike",
                     "NPClike",
                     "MES/AClike",
                     "NPClike",
                     "NPClike",
                     "MESlike",
                     "Alllike",
                     "NPClike", 
                     "AC/OPC/NPClike",
                     "GlutamatergicNeuron(L4)",
                     "NPClike",
                     "NPClike",
                     "NPClike",
                     "MES/OPClike",
                     "MES/AClike"),
                    
                    c("Alllike",
                      "Unclassified",
                      "Unclassified",
                      "MESlike",
                      "Unclassified",
                      "AC/OPC/NPClike",
                      "OPClike",
                      "MESlike",
                      "NPClike",
                      "Microglia"),
                    
                    c("Oligodendrocytes",
                     "OPC/NPClike",
                     "Alllike",
                     "Alllike",
                     "MESlike",
                     "OPC/NPClike",
                     "Microglia",
                     "NPClike",
                     "AClike", 
                     "OPC/NPClike",
                     "NPClike",
                     "MESlike",
                     "MESlike",
                     "MESlike"),
  
                    c("Oligodendrocytes",
                     "Microglia",
                     "Oligodendrocytes",
                     "Alllike",
                     "OPC/NPClike",
                     "AC/OPC/NPClike",
                     "MESlike",
                     "MESlike",
                     "MESlike", 
                     "OPC/NPClike",
                     "MES/AClike",
                     "MES/AClike",
                     "NPClike"))

```

Redo Heatmap with new cluster names:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:4){
pdf(paste('figures/KnownMarkersHeatmapClusterNames', as.character(count), '10x_Patient', names(Glioblastoma.10x)[i],'.pdf', sep = ''), width = 20, height = 10)
print(DoHeatmap(Glioblastoma.10x[[i]], features = c('MESlike', 'AClike', 'OPClike', 'NPClike', unlist(cell.markers)), slot = 'scale.data') + NoLegend())
dev.off()
print(DoHeatmap(Glioblastoma.10x[[i]], features = unlist(cell.markers), slot = 'scale.data') + NoLegend()) 
print(DoHeatmap(Glioblastoma.10x[[i]], features = c('MESlike', 'AClike', 'OPClike', 'NPClike'), slot = 'scale.data') + NoLegend()) 
print(DoHeatmap(Glioblastoma.10x[[i]], features = unlist(kriegstein.markers), slot = 'scale.data') + NoLegend())
}
```

Make a bar plot of cell proportions:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

tab = list()
for (i in 1:4){
  tab[[i]] = table(Idents(Glioblastoma.10x[[i]]))/sum(table(Idents(Glioblastoma.10x[[i]])))
}

# create a dataset
patient <- c(rep(names(Glioblastoma.10x)[[1]] , length(tab[[1]])), rep(names(Glioblastoma.10x)[[2]] , length(tab[[2]])),
             rep(names(Glioblastoma.10x)[[3]] , length(tab[[3]])), rep(names(Glioblastoma.10x)[[4]] , length(tab[[4]])))
celltype <- names(unlist(tab))
value <- unname(unlist(tab))
data <- data.frame(patient,celltype,value)
 
# Stacked + percent
pdf('figures/CellProportions_All_10X.pdf', width = 20, height = 10)
ggplot(data, aes(fill=patient, y=value, x=celltype)) + geom_bar(position=position_dodge(), stat="identity")
dev.off()
ggplot(data, aes(fill=patient, y=value, x=celltype)) + geom_bar(position=position_dodge(), stat="identity")

```


Integrate with smartSeq data (only for patient 46, because there is no matching data for other patients):

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

reference.list <- list(Glioblastoma.10x[[4]], Glioblastoma.SmartSeq[[1]])
features = SelectIntegrationFeatures(reference.list)
features = c(features, 'HOXB9')
glioblastoma.anchors <- FindIntegrationAnchors(object.list = reference.list, dims = 1:n_dimensions, anchor.features = features)
glioblastoma.integrated <- IntegrateData(anchorset = glioblastoma.anchors, dims = 1:n_dimensions, normalization.method = "LogNormalize")

```

Visualize integrated data:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

library(cowplot)
library(patchwork)
# switch to integrated assay. The variable features of this assay are automatically
# set during IntegrateData
DefaultAssay(glioblastoma.integrated) <- "integrated"

# Run the standard workflow for visualization and clustering
glioblastoma.integrated <- ScaleData(glioblastoma.integrated, verbose = FALSE)
glioblastoma.integrated <- RunPCA(glioblastoma.integrated, npcs = n_dimensions, verbose = FALSE)
glioblastoma.integrated <- RunUMAP(glioblastoma.integrated, reduction = "pca", dims = 1:n_dimensions)
p1 <- DimPlot(glioblastoma.integrated, reduction = "umap", group.by = "Technology")
p2 <- DimPlot(glioblastoma.integrated, reduction = "umap", label = TRUE, 
    repel = TRUE) + NoLegend()
pdf('figures/UMAPintegrated_Patient46.pdf', width = 20, height = 10)
print(p1 + p2)
dev.off()
p1+p2
```

Plot Hobxb9 expression:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

# Do not include unclassified cells in reference:

value = glioblastoma.integrated@assays$integrated@data['HOXB9',]
tech = glioblastoma.integrated$Technology

value_raw = c(reference.list[[1]]@assays$RNA@data['HOXB9',], reference.list[[2]]@assays$RNA@data['HOXB9',])

data_integrated <- data.frame(tech,value)

p = ggplot(data_integrated, aes(x=tech, y=value)) + geom_violin(trim=FALSE)
p + stat_summary(fun.data="mean_sdl", mult=1, geom="crossbar", width=0.2 ) + 
  geom_jitter(shape=16, position=position_jitter(0.3))

```


Now classify cells and make plot of cell proportions:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

# Do not include unclassified cells in reference:


modules = read.delim('/home/jovyan/HB_ZIK/1-s2.0-S0092867419306877-mmc2 (1)_SUBSET.csv', sep = ',')
features = unname(unlist(modules)[unlist(modules) != ''])

Reference = Glioblastoma.10x[[4]]

glioblastoma.anchors <- FindTransferAnchors(reference = Reference, query = Glioblastoma.SmartSeq[[1]], dims = 1:n_dimensions,
                                            features = features)
predictions <- TransferData(anchorset = glioblastoma.anchors, refdata = Idents(Reference), dims = 1:n_dimensions)
Glioblastoma.SmartSeq[[1]] <- AddMetaData(Glioblastoma.SmartSeq[[1]], metadata = predictions)

```

Also make plots of integrations scores:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

# Do not include unclassified cells in reference:

# hist(unname(Glioblastoma.SmartSeq[[1]]$prediction.score.max[Glioblastoma.SmartSeq[[1]]$predicted.id == '3: OPClike']),xlim = c(0,1), xlab = 'Classification Score', main = 'Histogram of OPClike Classification Scores')
# hist(unname(Glioblastoma.SmartSeq[[1]]$prediction.score.max[Glioblastoma.SmartSeq[[1]]$predicted.id == '4: OPClike']),xlim = c(0,1), xlab = 'Classification Score', main = 'Histogram of OPClike Classification Scores')
# hist(unname(Glioblastoma.SmartSeq[[1]]$prediction.score.max[Glioblastoma.SmartSeq[[1]]$predicted.id == '5: Microglia']),xlim = c(0,1), xlab = 'Classification Score', main = 'Histogram of Microglia Classification Scores')

```


Finally visualize cell proportions in 10x and SmartSeq data:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

# Do not include unclassified cells in reference:
tab = c(table(Idents(Reference))/sum(table(Idents(Reference))), table(Glioblastoma.SmartSeq[[1]]$predicted.id)/sum(table(Glioblastoma.SmartSeq[[1]]$predicted.id)))

# create a dataset:
tech <- c(rep("10X" , length(table(Idents(Reference)))) , rep("SmartSeq" , length(table(Glioblastoma.SmartSeq[[1]]$predicted.id))))
celltype <- names(tab)
value <- unname(tab)
data <- data.frame(tech,celltype,value)
 
# Stacked + percent:
pdf('figures/CellProportions_Patient46.pdf', width = 20, height = 10)
ggplot(data, aes(fill=tech, y=value, x=celltype)) + geom_bar(position=position_dodge(), stat="identity")
dev.off()
ggplot(data, aes(fill=tech, y=value, x=celltype)) + geom_bar(position=position_dodge(), stat="identity")
```

Now integrate all SmartSeq data with all 10x data

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:4){
  Glioblastoma.10x[[i]]@assays$RNA@scale.data = Glioblastoma.10x[[i]]@assays$RNA@scale.data[5:dim(Glioblastoma.10x[[i]]@assays$RNA@scale.data)[1], ]
}

```

First integrate all 10x data:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

glioblastoma.anchors <- FindIntegrationAnchors(object.list = Glioblastoma.10x, dims = 1:n_dimensions)
glioblastoma.10x.integrated <- IntegrateData(anchorset = glioblastoma.anchors, dims = 1:n_dimensions)

```

Then use the integrated data as a reference to classicy cell types:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

library(cowplot)
library(patchwork)
# switch to integrated assay. The variable features of this assay are automatically
# set during IntegrateData
DefaultAssay(glioblastoma.10x.integrated) <- "integrated"

# Run the standard workflow for visualization and clustering
glioblastoma.10x.integrated <- ScaleData(glioblastoma.10x.integrated, verbose = FALSE)
glioblastoma.10x.integrated <- RunPCA(glioblastoma.10x.integrated, npcs = n_dimensions, verbose = FALSE)
glioblastoma.10x.integrated <- RunUMAP(glioblastoma.10x.integrated, reduction = "pca", dims = 1:n_dimensions)
p1 <- DimPlot(glioblastoma.10x.integrated, reduction = "umap", group.by = "Technology")
p2 <- DimPlot(glioblastoma.10x.integrated, reduction = "umap", label = TRUE, 
    repel = TRUE) + NoLegend()
# pdf('figures/UMAPintegrated_Patient46.pdf', width = 20, height = 10)
# print(p1 + p2)
# dev.off()
p1+p2
```


Now use the integrated reference to classify cells:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

glioblastoma.anchors <- FindTransferAnchors(reference = glioblastoma.10x.integrated, query = Glioblastoma.SmartSeq[[2]], dims = 1:n_dimensions, features = features)
predictions <- TransferData(anchorset = glioblastoma.anchors, refdata = Idents(glioblastoma.10x.integrated), dims = 1:n_dimensions)
Glioblastoma.SmartSeq[[2]] <- AddMetaData(Glioblastoma.SmartSeq[[2]], metadata = predictions)

glioblastoma.anchors <- FindTransferAnchors(reference = glioblastoma.10x.integrated, query = Glioblastoma.SmartSeq[[3]], dims = 1:n_dimensions, features = features)
predictions <- TransferData(anchorset = glioblastoma.anchors, refdata = Idents(glioblastoma.10x.integrated), dims = 1:n_dimensions)
Glioblastoma.SmartSeq[[3]] <- AddMetaData(Glioblastoma.SmartSeq[[3]], metadata = predictions)

glioblastoma.anchors <- FindTransferAnchors(reference = glioblastoma.10x.integrated, query = Glioblastoma.SmartSeq[[4]], dims = 1:n_dimensions, features = features)
predictions <- TransferData(anchorset = glioblastoma.anchors, refdata = Idents(glioblastoma.10x.integrated), dims = 1:n_dimensions)
Glioblastoma.SmartSeq[[4]] <- AddMetaData(Glioblastoma.SmartSeq[[4]], metadata = predictions)

```

And again we make an overview of the classifications for each donor:

Finally visualize cell proportions in 10x and SmartSeq data:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

for (i in c(2,3,4)){
# Do not include unclassified cells in reference:
tab = c(table(Idents(glioblastoma.10x.integrated))/sum(table(Idents(glioblastoma.10x.integrated))), table(Glioblastoma.SmartSeq[[i]]$predicted.id)/sum(table(Glioblastoma.SmartSeq[[i]]$predicted.id)))

# Create a dataset
tech <- c(rep("10X" , length(table(Idents(glioblastoma.10x.integrated)))) , rep("SmartSeq" , length(table(Glioblastoma.SmartSeq[[i]]$predicted.id))))
celltype <- names(tab)
value <- unname(tab)
data <- data.frame(tech,celltype,value)
 
# Stacked + percent
pdf(paste('figures/CellProportions_Patient', names(Glioblastoma.SmartSeq)[i],'.pdf', sep = ''), width = 20, height = 10)
print(ggplot(data, aes(fill=tech, y=value, x=celltype)) + geom_bar(position=position_dodge(), stat="identity"))
dev.off()
print(ggplot(data, aes(fill=tech, y=value, x=celltype)) + geom_bar(position=position_dodge(), stat="identity"))

print(i)
print(table(Glioblastoma.SmartSeq[[i]]$predicted.id))

}

```

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

# Do not include unclassified cells in reference:
# hist(unname(Glioblastoma.SmartSeq[[2]]$prediction.score.max[Glioblastoma.SmartSeq[[2]]$predicted.id == '3: OPClike']),xlim = c(0,1), xlab = 'Classification Score', main = 'Histogram of OPClike Classification Scores')
# hist(unname(Glioblastoma.SmartSeq[[2]]$prediction.score.max[Glioblastoma.SmartSeq[[2]]$predicted.id == '4: OPClike']),xlim = c(0,1), xlab = 'Classification Score', main = 'Histogram of OPClike Classification Scores')
# hist(unname(Glioblastoma.SmartSeq[[1]]$prediction.score.max[Glioblastoma.SmartSeq[[1]]$predicted.id == '5: Microglia']),xlim = c(0,1), xlab = 'Classification Score', main = 'Histogram of Microglia Classification Scores')

```

To be sure we also cluster the SmartSeq data on its own:

Calculate principal components:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:4){
Glioblastoma.SmartSeq[[i]] <- RunPCA(Glioblastoma.SmartSeq[[i]], features = VariableFeatures(object = Glioblastoma.SmartSeq[[i]]))
print(Glioblastoma.SmartSeq[[i]][["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(Glioblastoma.SmartSeq[[i]], dims = 1:2, reduction = "pca")
DimPlot(Glioblastoma.SmartSeq[[i]], reduction = "pca")
DimHeatmap(Glioblastoma.SmartSeq[[i]], dims = 1, cells = round(dim(Glioblastoma.SmartSeq[[i]])[2])/2, balanced = TRUE)
DimHeatmap(Glioblastoma.SmartSeq[[i]], dims = 1:15, cells = round(dim(Glioblastoma.SmartSeq[[i]])[2])/2, balanced = TRUE)
}
```

This is the clustering step:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
n_dimensions = 15
for (i in 1:4){
Glioblastoma.SmartSeq[[i]] <- FindNeighbors(Glioblastoma.SmartSeq[[i]], dims = 1:n_dimensions)
Glioblastoma.SmartSeq[[i]] <- FindClusters(Glioblastoma.SmartSeq[[i]], resolution = 0.5)
head(Idents(Glioblastoma.SmartSeq[[i]]), 5)
}
```

Clusters roughly agree with visual seperation on a UMAP plot:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:4){
Idents(Glioblastoma.SmartSeq[[i]]) = Glioblastoma.SmartSeq[[i]]$predicted.id
Glioblastoma.SmartSeq[[i]] <- RunUMAP(Glioblastoma.SmartSeq[[i]], dims = 1:n_dimensions)

pdf(paste('figures/UMAPwithClusterLabel_SmartSeq_Patient', names(Glioblastoma.SmartSeq)[i],'.pdf', sep = ''), width = 20, height = 10)
print(DimPlot(Glioblastoma.SmartSeq[[i]], reduction = "umap", label = TRUE))
dev.off()
print(DimPlot(Glioblastoma.SmartSeq[[i]], reduction = "umap", label = TRUE))
}
```

Find and visualize top cluster markers:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
i = 1
markers_S <- FindAllMarkers(Glioblastoma.SmartSeq[[i]], only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
top = markers_S %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
pdf('figures/FeaturePlotUnbiasedClusterMarkers_SmartSeq_Patient46.pdf', width = 20, height = 10)
print(FeaturePlot(Glioblastoma.SmartSeq[[i]], features = top$gene))
dev.off()
print(FeaturePlot(Glioblastoma.SmartSeq[[i]], features = top$gene))
top10<- markers_S %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
pdf('figures/HeatmapUnbiasedClusterMarkers__SmartSeq_Patient46.pdf', width = 20, height = 10)
print(DoHeatmap(Glioblastoma.SmartSeq[[i]], features = top10$gene, cells = 1:dim(Glioblastoma.SmartSeq[[i]])[2] + NoLegend()) + 
    theme(text = element_text(size = 5)))
dev.off()
print(DoHeatmap(Glioblastoma.SmartSeq[[i]], features = top10$gene, cells = 1:dim(Glioblastoma.SmartSeq[[i]])[2] + NoLegend()) + 
    theme(text = element_text(size = 5)))
```

Vizualize a priori chosen markers of tumour subtypes:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:4){

print(i)
  
Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data = rbind(colSums(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data[rownames(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data) %in% c(modules['MES2'][modules['MES2'] != ''], modules['MES1'][modules['MES1'] != '']),]), Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data)
rownames(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data)[1] = 'MESlike'

Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data = rbind(colSums(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data[rownames(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data) %in% modules['AC'][modules['AC'] != ''],]), Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data)
rownames(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data)[1] = 'AClike'

Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data = rbind(colSums(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data[rownames(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data) %in% modules['OPC'][modules['OPC'] != ''],]), Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data)
rownames(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data)[1] = 'OPClike'

Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data = rbind(colSums(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data[rownames(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data) %in% c(modules['NPC2'][modules['NPC2'] != ''], modules['NPC1'][modules['NPC1'] != '']),]), Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data)
rownames(Glioblastoma.SmartSeq[[i]]@assays$RNA@scale.data)[1] = 'NPClike'

features = c('MESlike', 'AClike', 'OPClike', 'NPClike', unlist(cell.markers)) 
count = 0
for (f in list(1:10, 11:20,21:30,31:43)){
  count = count + 1
  pdf(paste('figures/KnownMarkersFeaturePlot', as.character(count), 'SmartSeq_Patient', names(Glioblastoma.SmartSeq)[i],'.pdf', sep = ''), width = 20, height = 10)
  print(FeaturePlot(Glioblastoma.SmartSeq[[i]], features = features[f], slot = 'scale.data'))
  dev.off()
  print(FeaturePlot(Glioblastoma.SmartSeq[[i]], features = features[f], slot = 'scale.data'))
}

pdf(paste('figures/KriegsteinMarkersFeaturePlot', '_SmartSeq_Patient', names(Glioblastoma.SmartSeq)[i],'.pdf', sep = ''), width = 20, height = 10)
print(FeaturePlot(Glioblastoma.SmartSeq[[i]], features = unlist(kriegstein.markers), slot = 'scale.data'))
dev.off()
print(FeaturePlot(Glioblastoma.SmartSeq[[i]], features = unlist(kriegstein.markers), slot = 'scale.data'))
}

```

Visualize a priori chosen markers in a heatmap:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:4){
pdf(paste('figures/KnownMarkersHeatmap', 'SmartSeq_Patient', names(Glioblastoma.SmartSeq)[i],'.pdf', sep = ''), width = 20, height = 10)
print(DoHeatmap(Glioblastoma.SmartSeq[[i]], features = c('MESlike', 'AClike', 'OPClike', 'NPClike', unlist(cell.markers)), slot = 'scale.data') + NoLegend())
dev.off()
print(DoHeatmap(Glioblastoma.SmartSeq[[i]], features = unlist(cell.markers), slot = 'scale.data') + NoLegend()) 
print(DoHeatmap(Glioblastoma.SmartSeq[[i]], features = c('MESlike', 'AClike', 'OPClike', 'NPClike'), slot = 'scale.data') + NoLegend()) 
print(DoHeatmap(Glioblastoma.SmartSeq[[i]], features = unlist(kriegstein.markers), slot = 'scale.data') + NoLegend())
}
```


```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

for (i in 1:length(Glioblastoma.SmartSeq)){
  Idents(Glioblastoma.SmartSeq[[i]]) = Glioblastoma.SmartSeq[[i]]$predicted.id
}

integration.list = unlist(list(glioblastoma.10x.integrated, Glioblastoma.SmartSeq))
glioblastomaAll.anchors <- FindIntegrationAnchors(object.list = integration.list, dims = 1:n_dimensions)
glioblastoma.All.integrated <- IntegrateData(anchorset = glioblastomaAll.anchors, dims = 1:n_dimensions)

```

Now also integrate SmartSeq data and 10X:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

library(cowplot)
library(patchwork)
# switch to integrated assay. The variable features of this assay are automatically
# set during IntegrateData
DefaultAssay(glioblastoma.10x.integrated) <- "integrated"

# Run the standard workflow for visualization and clustering
glioblastoma.All.integrated <- ScaleData(glioblastoma.All.integrated, verbose = FALSE)
glioblastoma.All.integrated <- RunPCA(glioblastoma.All.integrated, npcs = n_dimensions, verbose = FALSE)
glioblastoma.All.integrated <- RunUMAP(glioblastoma.All.integrated, reduction = "pca", dims = 1:n_dimensions)
p1 <- DimPlot(glioblastoma.All.integrated, reduction = "umap", group.by = "Technology")
p2 <- DimPlot(glioblastoma.All.integrated, reduction = "umap", label = TRUE, 
    repel = TRUE) + NoLegend()
# pdf('figures/UMAPintegrated_Patient46.pdf', width = 20, height = 10)
# print(p1 + p2)
# dev.off()
p1+p2
```

And make a UMAP plot of 10X and SmartSeq data side by side with the same axis and including cell type classifications:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

idents = Idents(glioblastoma.All.integrated)

idents[!idents %in% c('GlutamatergicNeuron(L4)', 'Unclassified', 'Microglia', 'Oligodendrocytes')] = 'Malignant'

Idents(glioblastoma.All.integrated) = idents

p1 <- DimPlot(glioblastoma.All.integrated[,glioblastoma.All.integrated$Technology == '10X' & glioblastoma.All.integrated$Patient != '46'], reduction = "umap", label = TRUE) + NoLegend()
p2 <- DimPlot(glioblastoma.All.integrated[,glioblastoma.All.integrated$Technology == 'SmartSeq' & glioblastoma.All.integrated$Patient != '46'], reduction = "umap", 
    repel = TRUE)
pdf('figures/UMAPintegratedAll_donorRest.pdf', width = 20, height = 10)
print(p1 + p2)
dev.off()
p1+p2

p1 <- DimPlot(glioblastoma.All.integrated[,glioblastoma.All.integrated$Technology == '10X' & glioblastoma.All.integrated$Patient == '46'], reduction = "umap", label = TRUE) + NoLegend()
p2 <- DimPlot(glioblastoma.All.integrated[,glioblastoma.All.integrated$Technology == 'SmartSeq' & glioblastoma.All.integrated$Patient == '46'], reduction = "umap", 
    repel = TRUE)
pdf('figures/UMAPintegratedAll_donor46.pdf', width = 20, height = 10)
print(p1 + p2)
dev.off()
p1+p2

```

Now make a heatmap showing expression of the main markers in 10x and SmartSeq data:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

for (i in 1:2){
  Glioblastoma.list[[i]] = NormalizeData(Glioblastoma.list[[i]], normalization.method = "LogNormalize", scale.factor = 10000)
  Glioblastoma.list[[i]] <- FindVariableFeatures(Glioblastoma.list[[i]], selection.method = "vst", nfeatures = 2000)
  Glioblastoma.list[[i]] = RunPCA(Glioblastoma.list[[i]], features = VariableFeatures(object = Glioblastoma.list[[i]]))
  Glioblastoma.list[[i]] <- ScaleData(Glioblastoma.list[[i]], features = all.genes)
  Glioblastoma.list[[i]] <- FindNeighbors(Glioblastoma.list[[i]], dims = 1:n_dimensions)
  Glioblastoma.list[[i]] <- FindClusters(Glioblastoma.list[[i]], resolution = 0.5)
}

plot_genes = modules[1:18,1:6]
labels = rep(c('transparent', 'black'), 54)
p1 = DoHeatmap(Glioblastoma.list[[1]][,Glioblastoma.list[[1]]$Patient != '46'], features = unlist(plot_genes), slot = 'scale.data', group.bar = FALSE, label = FALSE) + theme(axis.text.y = element_text(color = rev(x = labels))) + NoLegend()

p2 = DoHeatmap(Glioblastoma.list[[1]][,Glioblastoma.list[[1]]$Patient == '46'], features = unlist(plot_genes), slot = 'scale.data', group.bar = FALSE, label = FALSE) + theme(axis.text.y = element_text(color = rev(x = labels))) + NoLegend()

p3 = DoHeatmap(Glioblastoma.list[[2]][,sample(which(Glioblastoma.list[[2]]$Patient == '46'), 10000)], features = unlist(plot_genes), slot = 'scale.data', group.bar = FALSE, label = FALSE) + theme(axis.text.y = element_text(color = rev(x = labels))) + NoLegend()

p4 = DoHeatmap(Glioblastoma.list[[2]][,sample(which(Glioblastoma.list[[2]]$Patient != '46'), 10000)], features = unlist(plot_genes), slot = 'scale.data', group.bar = FALSE, label = FALSE) + theme(axis.text.y = element_text(color = rev(x = labels))) + NoLegend()

pdf('figures/Heatmap_DonorRest.pdf', width = 20, height = 10)
print(p4 + p1)
dev.off()

pdf('figures/HeatmapDonorRest_and_46.pdf', width = 20, height = 10)
print(p3 + p2)
dev.off()

p4 + p1 

p3 + p2

```
